---
title: "Nigeria Partner Deep Dive"
date: "November 15, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r dependencies, include=FALSE}
#load dependencies
library(tidyverse)
library(scales)
library(knitr)
library(kableExtra)
library(gridExtra)
library(ICPIutilities)
library(RearWindow)
```


```{r import, echo=FALSE}
#import,clean up names and add cumulative
  df_nga <- read_rds("~/ICPI/Data/MER_Structured_Dataset_PSNU_IM_FY17-18_20181115_v1_1_Nigeria.rds") %>% 
    rename_official() %>% 
    rw_gen_nn_target()
    
```

```{r graph_setup, echo=FALSE}
  #add palette
    color <- rw_addpalette()
  #establish thresholds
    threshold <- rw_addthresholds(df_nga)
```

```{r mech_filter, echo=FALSE}
#filter to look at just one mechanism
  df_mech <- df_nga %>% 
    filter(mechanismid == "14505") 

mech <- unique(df_mech$mechanismid)
mechname <- unique(df_mech$implementingmechanismname)
partner <- unique(df_mech$primepartner)

ind <- c("GEND_GBV", "HTS_TST", "HTS_TST_POS", "OVC_SERV", 
         "PMTCT_ART", "PMTCT_STAT", "TB_ART", "TB_STAT", 
         "TX_CURR", "TX_NEW", "TX_PVLS", "TX_RET")
```
## Deep dive for partner review in Nigeria.

The following visualizations depict FY18 progress for select PEPFAR MER indicators for `r partner`'s `r mech`: `r mechname`. 


## Overall achievement 

```{r achievement_data, echo=FALSE}

df_ach_mech <- df_mech %>% 
    filter(((indicator %in% ind) & 
            standardizeddisaggregate == "Total Numerator") |
         ((indicator %in% c("TB_ART", "TB_STAT", "TX_PVLS", "TX_RET")) &
           standardizeddisaggregate == "Total Denominator")) %>% 
  mutate(indicator = ifelse(numeratordenom == "D", paste0(indicator, "_D"), indicator)) %>% 
  group_by(mechanismid, indicator) %>% 
  summarise_at(vars(fy2018_targets, fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  filter(fy2018_targets != 0) %>% 
  mutate(fy2018ach = round(fy2018apr / fy2018_targets, 3)) %>% 
  arrange(fy2018ach) %>% 
  mutate(ach_color = case_when(fy2018ach < threshold["low"]  ~ color["ubuntu"],
                               fy2018ach < threshold["med"]  ~ color["amazon"],
                               TRUE                          ~ color["grullo"]))
  

df_ach_mech %>% 
  ggplot(aes(reorder(indicator, -fy2018ach), fy2018ach)) +
  geom_segment(aes(xend=indicator, y=0, yend=fy2018ach),
               color = df_ach_mech$ach_color,
               size = 2,
               na.rm = TRUE) +
  geom_point(color = df_ach_mech$ach_color,
             size = 6,
             na.rm = TRUE) +
  geom_text(aes(label = percent(fy2018ach)), 
                         hjust=-.5, vjust=.3, color = color["txtgray"],
                         na.rm = TRUE) +
  coord_flip() +
  labs(x = "", y = "", title = "Overall Achievement") + 
  rw_plot_theme() +
  theme(axis.text.x = ggplot2::element_blank())

rm(df_ach_mech)

```


## Comparing changes in cascade indicators between FY17 and FY18

```{r cascade_delta, echo=FALSE}

df_cascade <- df_mech %>% 
  filter(indicator %in% c("HTS_TST", "HTS_TST_POS", "TX_NEW", "TX_CURR"),
         standardizeddisaggregate == "Total Numerator") %>%
  select(-fy2019_targets) %>% 
  group_by(indicator) %>% 
  summarise_at(vars(matches("target|apr")), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  gather(ind, val, -indicator) %>% 
  mutate(pd = paste0("FY", str_sub(ind, 5L, 6L)),
         type = ifelse(str_detect(ind, "targets"), "TARGET", "RESULT")) %>% 
  select(-ind) %>% 
  spread(type, val) %>% 
  mutate(indicator = factor(indicator, levels = c("HTS_TST", "HTS_TST_POS", "TX_NEW", "TX_CURR")))
           

df_cascade %>% 
  ggplot(aes(pd, RESULT)) +
  geom_col() +
  geom_errorbar(aes(ymax=TARGET, ymin=TARGET), width=0.75, size =1, color= color["amazon"]) +
  scale_y_continuous(label = comma) +
  facet_wrap(indicator ~ ., scales = "free_y") +
  labs(x = "", y = "") +
  rw_plot_theme()


```

## Treatment cascade

``` {r cascade, echo=FALSE}
v_hts <- df_cascade %>% 
  filter(indicator == "HTS_TST",
         pd == "FY18") %>% 
  ggplot(aes(indicator, RESULT)) +
  geom_col() +
  geom_errorbar(aes(ymax=TARGET, ymin=TARGET), width=0.75, size =1, color= color["amazon"]) +
  scale_y_continuous(label = comma) +
  labs(x = "", y = "", title = "Cascade") +
  rw_plot_theme()

v_postx <- df_cascade %>% 
  filter(indicator %in% c("HTS_TST_POS", "TX_NEW"),
         pd == "FY18") %>% 
  ggplot(aes(indicator, RESULT)) +
  geom_col() +
  geom_errorbar(aes(ymax=TARGET, ymin=TARGET), width=0.75, size =1, color= color["amazon"]) +
  scale_y_continuous(label = comma) +
  labs(x = "", y = "", title = " ") +
  rw_plot_theme()

grid.arrange(v_hts, v_postx, ncol = 2)
  rm(v_hts, v_postx, df_cascade)
```

## Cascade table

```{r cascade_table, echo=FALSE}

df_mech %>% 
  filter(indicator %in% c("HTS_TST", "HTS_TST_POS", "TX_NEW", "TX_CURR"),
         standardizeddisaggregate == "Total Numerator") %>%
  group_by(indicator) %>% 
  summarise_at(vars(matches("fy2018")), sum, na.rm = TRUE) %>% 
  ungroup() %>%
  gather(pd, val, -indicator) %>% 
  mutate(pd = str_remove(pd, "fy2018_|fy2018") %>% toupper(.)) %>% 
  spread(indicator, val) %>% 
  mutate(pd = factor(pd, levels = c("Q1", "Q2", "Q3", "Q4", "APR", "TARGETS"))) %>% 
  arrange(pd) %>% 
  mutate(Linkage = percent(TX_NEW / HTS_TST_POS, accuracy = 2),
         Positivity = percent(HTS_TST_POS / HTS_TST,accuracy = 2)) %>% 
  kable(format.args = list(big.mark = ",", zero.print = FALSE)) %>% 
  #add_header_above(c(" " = 1, "FY18 Q1" = 3, "Q2" = 3, "Q3" = 3, "Q4" = 3)) %>% 
  kable_styling()
```

## HTS achievement & positivity by state

```{r hts_state, echo=FALSE}

df_hts_state <- df_mech %>% 
  filter(indicator == "HTS_TST",
         standardizeddisaggregate == "Total Numerator") %>%
  group_by(snu1) %>% 
  summarize_at(vars(fy2018apr, fy2018_targets), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  mutate(fy2018ach = round(fy2018apr/fy2018_targets, 3))

v_test <- df_hts_state %>% 
  ggplot(aes(reorder(snu1, fy2018apr), fy2018apr)) +
  geom_col() +
  geom_errorbar(aes(ymax=fy2018_targets, ymin=fy2018_targets), width=0.75, size =1, color= color["amazon"]) +
  scale_y_continuous(label = comma) +
  coord_flip() +
  labs(x= "", y = "testing volume, FY18", title = "Testing volume by state") +
  rw_plot_theme() #+
  #theme(panel.grid.major = element_line(color=color["lgray"]))

df_yield <- df_mech %>% 
  filter(indicator %in% c("HTS_TST", "HTS_TST_POS"),
         standardizeddisaggregate == "Total Numerator") %>%
  group_by(snu1, indicator) %>% 
  summarize_at(vars(fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  spread(indicator, fy2018apr) %>% 
  filter(!is.na(HTS_TST)) %>% 
  mutate(positivity = HTS_TST_POS / HTS_TST)

v_yield <- df_yield %>% 
  ggplot(aes(reorder(snu1, HTS_TST), positivity)) +
  geom_col() +
  geom_text(aes(y = 0, label = percent(positivity, accuracy = 1)), hjust = -.2) +
  scale_y_continuous(label = percent) +
  coord_flip() +
  labs(x = " ", y = "", title = "Testing Positivity") +
  rw_plot_theme() +
  theme(axis.text.x = element_text(color = "white"),
        axis.text.y = element_text(color = "white"))
  
grid.arrange(v_test, v_yield, ncol = 2)
rm(df_hts_state, v_test, v_yield, df_yield)
```

## Testing to treatment (proxy linkage)

```{r htstx, echo=FALSE}
df_htstx <- df_mech %>% 
  filter(indicator %in% c("HTS_TST_POS", "TX_NEW"),
         standardizeddisaggregate == "Total Numerator") %>%
  group_by(snu1, indicator) %>% 
  summarise_at(vars(fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  filter(fy2018apr != 0)

lvls <- df_htstx %>% 
  filter(indicator == "HTS_TST_POS") %>% 
  arrange(desc(fy2018apr)) %>% 
  pull(snu1) %>% 
  fct_inorder()

df_htstx %>%
  mutate(snu1 = factor(snu1, lvls)) %>% 
  ggplot(aes(indicator, fy2018apr)) +
  geom_col() +
  geom_text(y = 0, aes(label = comma(fy2018apr)), vjust = -.5) +
  facet_wrap(snu1 ~ .) +
  #scale_y_continuous(label = comma) +
  labs (x = "", y = "") +
  rw_plot_theme() +
  theme(axis.text.y = element_blank())

rm(lvls)  
```

## Testing to treatment (proxy linkage) table

```{r htstx table, echo=FALSE}
df_htstx %>% 
  spread(indicator, fy2018apr) %>% 
  mutate(`Proxy Linkage (%)` = round(TX_NEW / HTS_TST_POS, 2) * 100) %>% 
  arrange(desc(HTS_TST_POS)) %>% 
  rename(State = snu1) %>% 
  kable(format.args = list(big.mark = ",", zero.print = FALSE)) %>% 
  kable_styling()
```


```{r, echo=FALSE}
df_mod <- df_mech %>% 
  filter(indicator %in% c("HTS_TST", "HTS_TST_POS"),
         standardizeddisaggregate %in% 
           c("Modality/Age/Sex/Result", "Modality/Age Aggregated/Sex/Result")) %>% 
  group_by(modality, indicator) %>% 
  summarize_at(vars(fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  spread(indicator, fy2018apr) %>% 
  filter(HTS_TST != 0) %>% 
  mutate(`Positivity (%)` = round(HTS_TST_POS / HTS_TST, 3)*100) %>% 
  arrange(HTS_TST) %>% 
  mutate(modality = as_factor(modality)) %>% 
  gather(indicator, val, -modality)

df_mod %>% 
  ggplot(aes(modality, val)) +
  geom_col() +
  scale_y_continuous(labels = comma) +
  coord_flip() +
  facet_wrap(indicator ~ ., scales = "free_x") +
  labs(title = "Testing by Modality", x = "", y = "volume")
```

#HTS modalities table

```{r hts_mods_tbl, echo=FALSE}

# df_mech %>% 
#   filter(indicator %in% c("HTS_TST","HTS_TST_NEG", "HTS_TST_POS"),
#          standardizeddisaggregate %in% 
#            c("Modality/Age/Sex/Result", "Modality/Age Aggregated/Sex/Result")) %>% 
#   group_by(modality, indicator) %>% 
#   summarize_at(vars(starts_with("fy2018q")), sum, na.rm = TRUE) %>% 
#   ungroup() %>% 
#   gather(pd, val, -modality, -indicator) %>%
#   mutate(pd = str_remove(pd, "20") %>% toupper()) %>% 
#   spread(indicator, val) %>% 
#   mutate(YIELD = round(HTS_TST_POS / HTS_TST, 3) *100) %>% 
#   filter(!is.nan(YIELD)) %>% 
#   select(-HTS_TST_NEG) %>% 
#   gather(indicator, val, -modality, -pd) %>%
#   mutate(indicator = case_when(indicator == "HTS_TST" ~ "Total",
#                                indicator == "HTS_TST_POS" ~ "Pos.",
#                                TRUE                       ~ "Yield")) %>% 
#   arrange(modality, pd, indicator) %>% 
#   unite(ind, c("pd", "indicator")) %>% 
#   spread(ind, val, fill = 0) %>% 
#   kable(format.args = list(big.mark = ",", zero.print = FALSE),
#         col.names = c("", rep(c("Pos", "Total", "Yield (%)"), 4))) %>% 
#   add_header_above(c(" " = 1, "FY18 Q1" = 3, "Q2" = 3, "Q3" = 3, "Q4" = 3)) %>% 
#   kable_styling()

#apr
df_mech %>% 
  filter(indicator %in% c("HTS_TST","HTS_TST_NEG", "HTS_TST_POS"),
         standardizeddisaggregate %in% 
           c("Modality/Age/Sex/Result", "Modality/Age Aggregated/Sex/Result")) %>% 
  group_by(modality, indicator) %>% 
  summarize_at(vars(fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  spread(indicator, fy2018apr) %>% 
  mutate(Yield = percent(HTS_TST_POS / HTS_TST, accuracy = 1)) %>% 
  filter(Yield != "NaN%") %>% 
  select(-HTS_TST_NEG) %>% 
  rename(Total = HTS_TST,
         `Pos.` = HTS_TST_POS) %>% 
  arrange(desc(Total)) %>% 
  kable(format.args = list(big.mark = ",", zero.print = FALSE)) %>% 
  kable_styling()

```

```{r hts_mod_sex, echo=FALSE}

df_hts_mod <- df_mech %>% 
  filter(indicator == "HTS_TST_POS",
         standardizeddisaggregate %in% 
           c("Modality/Age/Sex/Result", "Modality/Age Aggregated/Sex/Result")) %>% 
  group_by(modality) %>% 
  summarize_at(vars(fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  filter(fy2018apr != 0)

df_hts_mod %>% 
  ggplot(aes(reorder(modality, fy2018apr), fy2018apr)) +
  geom_col() +
  coord_flip() +
  labs(title = "Positive Testing by Modality, FY18",x = "", y = "positive tests, FY18") +
  scale_y_continuous(label = comma) +
  rw_plot_theme()

df_hts_sex <- df_mech %>% 
  filter(indicator == "HTS_TST_POS",
         standardizeddisaggregate %in% 
           c("Modality/Age/Sex/Result", "Modality/Age Aggregated/Sex/Result")) %>% 
  group_by(modality, sex) %>% 
  summarize_at(vars(fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  filter(fy2018apr != 0) %>% 
  group_by(modality) %>% 
  mutate(mod_total = sum(fy2018apr)) %>% 
  ungroup() %>% 
  mutate(sex = ifelse(sex == "Unknown Sex", "Unknown", sex))

df_hts_sex %>% 
  filter(sex != "Unknown") %>% 
  ggplot(aes(reorder(modality, mod_total), fy2018apr)) +
  geom_col() +
  coord_flip() +
  facet_grid(~ sex) +
  labs(x = "", y = "positive tests, cumulative") +
  scale_y_continuous(label = comma) +
  rw_plot_theme()

df_hts_age <- df_mech %>% 
  filter(indicator %in% c("HTS_TST","HTS_TST_POS"),
         standardizeddisaggregate %in% 
           c("Modality/Age/Sex/Result", "Modality/Age Aggregated/Sex/Result"),
         agesemifine != "Unknown Age") %>% 
  group_by(indicator, agesemifine) %>% 
  summarize_at(vars(fy2018apr), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  spread(indicator, fy2018apr) %>% 
  mutate(`Positivity (%)` = round(HTS_TST_POS / HTS_TST, 3) * 100) %>% 
  filter(HTS_TST != 0) %>%
  gather(indicator, fy2018apr, -agesemifine) %>% 
  dplyr::arrange(desc(agesemifine)) %>%
  mutate(agesemifine = forcats::as_factor(agesemifine))

df_hts_age %>% 
  ggplot(aes(agesemifine, fy2018apr)) +
  geom_col() +
  coord_flip() +
  facet_wrap(~ indicator, scales = "free_x") +
  labs(x = "", y = "test volume") +
  scale_y_continuous(label = comma) +
  rw_plot_theme() 


df_hts_age %>% 
  ggplot(aes(agesemifine, fy2018apr, color = indicator)) +
  geom_point(size = 6) +
  coord_flip() +
  labs(x = "", y = "test volume") +
  scale_y_continuous(label = comma) +
  rw_plot_theme()  

```

## Treatment

```{r tx_table, echo=FALSE}

df_mech %>% 
  filter(indicator %in% c("TX_NEW", "TX_CURR", "TX_NET_NEW"),
         standardizeddisaggregate == "Total Numerator") %>% 
  group_by(indicator) %>% 
  summarize_at(vars(starts_with("fy2018q"), fy2018apr, fy2018_targets), sum, na.rm = TRUE) %>% 
  ungroup() %>% 
  gather(pd, val, starts_with("fy2018")) %>% 
  spread(indicator, val) %>% 
  mutate(pd = factor(pd, 
                     levels = c("fy2018q1", "fy2018q2", "fy2018q3", "fy2018q4", "fy2018apr", "fy2018_targets"))) %>% 
  arrange(pd) %>% 
  mutate(pd = str_remove_all(pd, "20") %>% toupper(.),
         pd = str_replace(pd, "_", " "),
         pd = str_replace(pd, "A", " A")) %>% 
  kable(format.args = list(big.mark = ",", zero.print = FALSE)) %>%
  kable_styling()

```

## TX_NET_NEW

```{r txnetnew, echo=FALSE}

df_nn <- df_mech %>% 
  filter(indicator %in% c("TX_CURR", "TX_NET_NEW"),
         standardizeddisaggregate == "Total Numerator") %>% 
  group_by(indicator) %>% 
  summarise_at(vars(contains("q")), sum, na.rm = TRUE) %>% 
  gather(pd, val, -indicator) %>% 
  spread(indicator, val) %>% 
  mutate(tx_curr_gr = ifelse(TX_NET_NEW > 0, TX_CURR - TX_NET_NEW, TX_CURR),
         tx_nn_gr = ifelse(TX_NET_NEW > 0, TX_NET_NEW, 0)) %>% 
  select(-TX_CURR, -TX_NET_NEW) %>% 
  gather(indicator, val, -pd) %>% 
  mutate(pd = str_remove(pd, "20") %>% toupper(.),
         indicator = case_when(indicator == "tx_nn_gr" ~ "TX_NET_NEW",
                               TRUE                    ~ "TX_CURR"),
         indicator = factor(indicator, levels = c("TX_NET_NEW","TX_CURR")))
  
df_nn %>% 
  ggplot(aes(pd, fill = indicator)) + 
  geom_bar(aes(weight = val), 
           position = "stack") +
  scale_y_continuous(label = comma) +
  labs(x = "", y = "") +
  rw_plot_theme()
  
  
```



