---
title: "Low Yield Sites"
author: "Aaron Chafetz"
date: "December 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(scales)

#import
folderpath <- "C:/Users/achafetz/Documents/GitHub/DeepDive/Output"
tx_sites <- read_csv(file.path(folderpath, "lowyield_tx.csv"))
pmtct_sites <- read_csv(file.path(folderpath, "lowyield_pmtct.csv"))


main_blue <- "#6585CF"
main_gray <- "#4D4D4D"
comp_gray <- "#808080"
comp_teal <- "#6BB1C9"
comp_purp <- "#7E6BC9"
goal_gold <- "#CEB966"

```

## Overview

This report is a quick look into the volume of sites that are low yield, ie have 10 or less patients on treatment or have a positivity of zero for PMTCT_STAT.

## TX_CURR

```{r tx_curr}
#how many sites are reporting TX_CURR (exlusive of reporting 0)?
tot_tx <- tx_sites %>%
  distinct(sitename) %>% 
  nrow()

#how many sites are reporting <=10 TX_CURR?
low_tx <- tx_sites %>% 
  filter(tx_curr_fy18apr <= 10) %>% 
  nrow()

#how many USAID sites are reporting TX_CURR (exlusive of reporting 0)?
usaid_tx <- tx_sites %>%
  filter(fundingagency == "USAID") %>% 
  distinct(sitename) %>% 
  nrow()

#how many USAID sites are reporting <10 TX_CURR?
usaid_low_tx <- tx_sites %>% 
  filter(fundingagency == "USAID",
         tx_curr_fy18apr <= 10) %>% 
  nrow()

```

Looking first at TX_CURR, were `r tot_tx` sites reporting on TX_CURR at FY2018 APR in Nigeria. USAID is in `r usaid_tx` of these facilities. Of all the sites reporting on TX_CURR, `r low_tx` have 10 or less patients on treatment. USAID is in  `r percent(usaid_low_tx/low_tx, accuracy = 1)` of those low sites.

Nearly two-thirds of USAID's sites are in four states - Rivers, Akwa Ibom, Lagos, and Niger.

```{r}

snu_order <- tx_sites %>% 
  filter(snu1 != "_Military Nigeria") %>% 
  count(snu1) %>% 
  arrange(n) %>% 
  pull(snu1)

tx_sites %>% 
  filter(fundingagency == "USAID",
         tx_curr_fy18apr <= 10) %>% 
  count(snu1, sort = TRUE) %>% 
  mutate(cum = cumsum(n)/sum(n))


tx_snu <- tx_sites %>% 
  filter(snu1 != "_Military Nigeria") %>% 
  mutate(reporting = ifelse(tx_curr_fy18apr > 10, "Reporting", "None"),
         snu1 = factor(snu1, levels = snu_order),
         reporting = factor(reporting, levels = c("Reporting", "None")),
         rep_col = ifelse(reporting == "Reporting", comp_gray, main_blue)) %>% 
  count(snu1, fundingagency, reporting, rep_col) %>% 
  arrange(snu1)
  
tx_snu %>% 
  ggplot(aes(snu1, n, fill = reporting)) +
  geom_col() +
  scale_fill_manual(values=alpha(c(main_gray, main_gray))) +
  coord_flip() +
  labs(x = "", y = "number of sites") +
  theme(legend.position = "none")

tx_snu %>% 
  ggplot(aes(snu1, n, fill = reporting)) +
  geom_col(fill = tx_snu$rep_col) +
  #scale_fill_manual(values=alpha(c(comp_gray, main_blue))) +
  coord_flip() +
  labs(x = "", y = "number of sites") +
  theme(legend.position = "none")


tx_sites %>% 
  filter(snu1 != "_Military Nigeria",
         tx_curr_fy18apr < 10) %>% 
  count(snu1, fundingagency, sort = TRUE) %>% 
  ggplot(aes(reorder(snu1, n), n, fill = fundingagency)) +
  geom_col() + 
  coord_flip() +
  labs(x = "", y = "# of sites") +
  theme(axis.ticks = element_blank()) +
  coord_flip() +
  labs(x = "", y = "# of sites") +
  theme(axis.ticks = element_blank())


tx_sites %>% 
  filter(snu1 != "_Military Nigeria") %>% 
  count(snu1, fundingagency, sort = TRUE) %>% 
  ggplot(aes(reorder(snu1, n), n)) +
  geom_col() + 
  coord_flip() +
  labs(x = "", y = "# of sites") +
  theme(axis.ticks = element_blank())
  
  

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
