---
title: "Prevalence in Nigeria"
author: "Aaron Chafetz"
date: "Updated: Jan 15, 2019"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.width=12, fig.height=6)

#dependencies
library(tidyverse)
library(readxl)
library(scales)

#colors

main_blue <- "#6585CF"
main_gray <- "#4D4D4D"
comp_gray <- "#989898" #"#808080"
comp_teal <- "#6BB1C9"
comp_purp <- "#7E6BC9"
goal_gold <- "#CEB966"

```

```{r import}

#import
folderpath <- "C:/Users/achafetz/Documents/GitHub/DeepDive/Output"
prev <- read_rds(file.path(folderpath, "prev_surveys.rds"))
burden <- read_rds(file.path(folderpath, "nga_buden.rds"))
datapack <- read_rds(file.path(folderpath, "datapack.rds"))
cascade <- read_rds(file.path(folderpath, "cascade.rds"))
tx_curr <- read_rds(file.path(folderpath, "tx_curr.rds"))
naiis <- read_rds(file.path(folderpath, "naiis.rds"))

```

## NAIIS Notes

- Updated from December Retreat
- Final results will be released in March



**NOTE: NAIIS data is embargoed and unvalidated at this point**

## Discussion Points to consider

- What can we expect post NAIIS based on other country experiences? 
- How to interpret the current Nigeria NAIIS results? What do they mean and what don't they mean? 
- What should be USAID’s focus? Geographic, priority populations, population characteristics, service delivery models, targets, budgets?

```{r calc_decline}
n_decrease <- prev %>% 
  select(-lab_state) %>% 
  filter(survey != "Spectrum2016") %>% 
  spread(survey, prevalence) %>% 
  filter(Spectrum2018 > NAIIS2018) %>% 
  nrow()

n_states <- unique(prev$state) %>% length()
```
## The initial results show a decline in `r n_decrease` of the `r n_states` states compared with recent Spectrum prevalence.

```{r prev}
  
  rm(n_decrease, n_states)

  viz_prev <- prev %>% 
    select(-lab_state) %>% 
    spread(survey, prevalence) %>% 
    mutate(change = Spectrum2018 - NAIIS2018) %>% 
    arrange(desc(change)) %>% 
    mutate(state = as_factor(state)) 
  
  grwth_states <- viz_prev %>%
    mutate(state = as.character(state)) %>% 
    filter(change < 0) %>% 
    pull(state)
  
  viz_prev <- viz_prev %>%
    select(-change) %>% 
    gather(survey, prevalence, -state) %>% 
    mutate(col = case_when(state %in% grwth_states ~ comp_gray,
                           TRUE                    ~ main_blue),
           survey = str_replace(survey, "pectrum20|AIIS20", "'"),
           survey = factor(survey, levels = c("S'16", "S'18", "N'18")))

  viz_prev %>% 
      ggplot(aes(survey, prevalence, group = state)) +
      geom_path(size = 2, color = viz_prev$col) +
      geom_point(size = 6, pch=21, color = "white", fill = viz_prev$col, stroke = 2) +
      scale_y_continuous(labels=function(x) paste0(x *100,"%"),
                         breaks = seq(0, .10, .05)) +
      labs(x = "", caption = "S = Spectrum, N = NAIIS") +
      facet_wrap(. ~ state) +
      expand_limits(y = .1) +
      theme_bw() +
      theme(legend.position = "none",
            text = element_text(size = 18),
            axis.ticks = element_blank(),
            axis.title = element_text(color = main_gray),
            panel.border = element_blank(),
            axis.line =  element_blank(),
            strip.background = element_blank(),
            plot.caption = element_text(color = main_gray)) 
  
  rm(viz_prev, grwth_states)
```

## Here is where Nigeria was with PLHIV (Data Pack COP 18)

```{r plhiv}

  viz_plhiv <- burden %>% 
    select(state, plhiv_old, plhiv_est) %>%
    arrange(plhiv_old) %>% 
    mutate(state = as_factor(state)) %>% 
    gather(pd, val, -state) %>% 
    mutate(lab = case_when(state == "Anambra" & pd == "plhiv_old" ~ "Old",
                           state == "Anambra" & pd == "plhiv_est" ~ "NAIIS",
                          TRUE                                 ~ as.character(NA)))

  viz_plhiv_i <- viz_plhiv %>%
    filter(pd == "plhiv_old")
  
  viz_plhiv_i %>% 
    ggplot(aes(state, val)) +
    geom_point(color = ifelse(viz_plhiv_i$pd == "plhiv_old", comp_gray, main_blue),
               size = 6) +
    geom_text(aes(label = lab), 
              hjust = .5, vjust = -1, 
              check_overlap = TRUE,
              size = 5,
              color = ifelse(viz_plhiv_i$pd == "plhiv_old", comp_gray, main_blue),
              na.rm = TRUE) +
    scale_y_continuous(limits = c(0, 305000), label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(y = "PLHIV estimate", x = "") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 18),
          axis.ticks = element_blank(),
          axis.title = element_text(color = main_gray),
          panel.border = element_blank(),
          axis.line = element_line(color = main_gray))
  
  rm(viz_plhiv_i)

```

```{r calc_states}
  n_states <- nrow(burden)
  
  n_decrease <- burden %>% 
   filter(plhiv_est < plhiv_old) %>% 
   nrow()
  
  n_increase <- n_states - n_decrease
```

## ... and now NAIIS projects lower PLHIV in `r n_decrease` of the `r n_states` initial states.

```{r plhiv_updated}
  
  viz_plhiv %>% 
    ggplot(aes(state, val)) +
    geom_path(color = main_blue,
              size = 1, na.rm = TRUE) +
    geom_point(color = ifelse(viz_plhiv$pd == "plhiv_old", comp_gray, main_blue),
               size = 6, na.rm = TRUE) +
    geom_text(aes(label = lab), 
              hjust = .5, vjust = -1, 
              check_overlap = TRUE,
              size = 5,
              color = ifelse(viz_plhiv$pd == "plhiv_old", comp_gray, main_blue),
              na.rm = TRUE) +
    scale_y_continuous(limits = c(0, 305000), label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(y = "PLHIV estimate", x = "") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 18),
          axis.ticks = element_blank(),
          axis.title = element_text(color = main_gray),
          panel.border = element_blank(),
          axis.line = element_line(color = main_gray))
  
  rm(viz_plhiv)
```

## ...which means `r n_increase` states saw an increase in PLHIV between the surveys.

```{r plhiv_inc}
  
  rm(n_states, n_decrease)

  viz_plhiv_inc <- burden %>% 
    select(state, plhiv_old, plhiv_est) %>%
    arrange(plhiv_old) %>% 
    mutate(state = as_factor(state),
           increase = plhiv_est > plhiv_old) %>% 
    gather(pd, val, -state, -increase) %>% 
    mutate(lab = case_when(state == "Anambra" & pd == "plhiv_old" ~ "Old",
                           state == "Anambra" & pd == "plhiv_est" ~ "NAIIS",
                           TRUE                                 ~ as.character(NA)),
           val = ifelse(increase == TRUE, val, NA))
  
  viz_plhiv_inc %>% 
    ggplot(aes(state, val)) +
    geom_path(color = main_blue,
              size = 1, na.rm = TRUE) +
    geom_point(color = ifelse(viz_plhiv_inc$pd == "plhiv_old", comp_gray, main_blue),
               size = 6, na.rm = TRUE) +
    geom_text(aes(label = lab), 
              hjust = .5, vjust = -1, 
              check_overlap = TRUE,
              size = 5,
              color = ifelse(viz_plhiv_inc$pd == "plhiv_old", comp_gray, main_blue),
              na.rm = TRUE) +
    scale_y_continuous(limits = c(0, 305000), label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(y = "PLHIV estimate", x = "") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 18),
          axis.ticks = element_blank(),
          axis.title = element_text(color = main_gray),
          panel.border = element_blank(),
          axis.line = element_line(color = main_gray))
  
  rm(viz_plhiv_inc)
```

## Let's just look at the new PLHIV by state

```{r new_plhiv}
  #unmet need
  viz_burden <- burden %>% 
    arrange(plhiv_old) %>% 
    select(state, TX_CURR, unmet_burden) %>% 
    #arrange(unmet_burden) %>% 
    #filter(!is.na(TX_CURR)) %>% 
    mutate(state = as_factor(state)) %>% 
    gather(type, val, -state) %>% 
    arrange(state, type) %>% 
    mutate(col_type = ifelse(type == "TX_CURR", comp_gray, comp_teal),
           type = ifelse(type == "unmet_burden", "Unmet Need", type),
           type = as_factor(type),
           val = ifelse(is.na(val), 0, val)) 
  
  #function to graph plhiv/burden
  grph_burden <- function(fill_col){
             viz_burden %>% 
               ggplot(aes(state, val)) +
               geom_col(fill = fill_col) +
               scale_y_continuous(limits = c(0, 305000), label = comma) +
               coord_flip() +
               labs(x = "", y = "PLHIV") +
               theme_bw() +
               theme(text = element_text(size = 18),
                     axis.title = element_text(color = main_gray),
                     axis.ticks = element_blank(),
                     panel.border = element_blank(),
                     plot.caption = element_text(color = main_gray))
           }
  
  grph_burden(main_blue)

```

## We should reorder our visual on the new PLHIV estimates.

```{r new_plhiv_ordered}
  #Reordered PLHIV
  state_order_new <- burden %>% 
    arrange(plhiv_est) %>% 
    pull(state)
  
  viz_burden <- viz_burden %>% 
    mutate(state = factor(state, state_order_new)) %>% 
    arrange(state)
  
  grph_burden(main_blue)
  
  rm(state_order_new)
  
```


## Breaking down PLHIV, we can see Nigeria still has a large unmet need

```{r burden}
#unmet need
 grph_burden(viz_burden$col_type)

```


In the above graph, gray bars depict the amount of treatment coverage (FY18 APR MER TX_CURR) and teal shows our unmet need/burden.

## ... and shows Nigeria still has a ways to go before reaching the second 90.

```{r burden_share}

  viz_burden %>% 
      ggplot(aes(state, val, fill = type)) +
      geom_bar(position = "fill",stat = "identity", fill = viz_burden$col_type) +
      geom_hline(linetype = "dashed", yintercept = .81, color = goal_gold, size = 1) +
      scale_y_continuous(label = percent) +
      coord_flip() +
      labs(x = "", y = "Share of PLHIV", caption = " ") +
      theme_bw() +
      theme(text = element_text(size = 18),
            axis.title = element_text(color = main_gray),
            axis.ticks = element_blank(),
            panel.border = element_blank(),
            plot.caption = element_text(color = main_gray))
  
  rm(viz_burden, grph_burden)
```

In the above graph, gray bars depict the amount of treatment coverage (FY18 APR MER TX_CURR) and teal shows our unmet need/burden.

## It doesn't appear that our targets heavily align with this unmet need.

```{r burden_mer}
  viz_ind <- burden %>% 
    arrange(unmet_burden) %>% 
    mutate(state = as_factor(state)) %>% 
    select(state, `Unmet Burden` = unmet_burden, TX_NEW, HTS_TST) %>% 
    gather(type, val, -state) %>% 
    filter(!is.na(val)) %>% 
    mutate(type = factor(type, c("Unmet Burden", "HTS_TST", "TX_NEW")),
           val = val / 1000)

  viz_ind %>% 
    ggplot(aes(state, val)) +
    geom_col(fill = ifelse(viz_ind$type == "Unmet Burden", comp_teal, comp_purp)) +
    coord_flip() +
    facet_grid(. ~ type, scales = "free_x") +
    scale_y_continuous(label = comma) +
    labs(x = "", y = "", caption = "values in 1,000's") +
    theme_bw() +
    theme(text = element_text(size = 18),
          axis.title = element_text(color = main_gray),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(fill = "white", color = "white"),
          strip.text = element_text(color = main_gray))
  
  rm(viz_ind)
```


## The cascade shows low 1st 90, but sucessful 2nd and 3rd 90.

```{r cascade, fig.width= 16}

  datapack_sex <- datapack %>% 
    select(-plhiv_old) %>% 
    gather(sex, pop, -state) %>% 
    mutate(sex = case_when(sex == "pop_num_f" ~ "female",
                           sex == "pop_num_m" ~ "male",
                           TRUE               ~ "total"))
  
  cascade_est <- left_join(naiis, datapack_sex, by = c("state", "sex"))
  
  # viz_cas <- cascade_est %>%
  #   mutate(PLHIV = pop * `Prev 15+`,
  #          Diagnosed = PLHIV * Diagnosed,
  #          `On Treatment` = Diagnosed * `On Treatment`,
  #          `Virally Suppressed` = `On Treatment` * `Virally Suppressed`) %>% 
  #   select(-`Prev 15+`, -pop, -PLHIV) %>% 
  #   gather(type, val, -state, -sex)
  # 
  
  viz_cas <- cascade_est %>%
    mutate(`On Treatment` = Diagnosed * `On Treatment`,
           `Virally Suppressed` = `On Treatment` * `Virally Suppressed`) %>% 
    select(-`Prev 15+`, -pop) %>% 
    gather(type, val, -state, -sex) %>% 
    mutate(type = ifelse(type == "On Treatment", "On Tx", type),
           type = case_when(type == "Diagnosed" ~ "1st 90",
                            type == "On Tx"     ~ "2nd 90",
                            type == "Virally Suppressed" ~ "3rd 90"),
           sex = str_to_title(sex),
           sex = factor(sex, levels = c("Total", "Female", "Male")),
           val = ifelse(is.na(val), 0, val))
  
  #order
  state_order <- viz_cas %>% 
    filter(sex == "Total", type == "1st 90") %>% 
    arrange(desc(val)) %>% 
    pull(state)
  
  viz_cas <- viz_cas %>% 
    mutate(state = factor(state, levels = state_order))
  
 grph_cas <- function(sex_type, color){
  
  viz_cas %>% 
    filter(sex == sex_type) %>% 
    ggplot(aes(type, val, fill = sex)) +
    geom_col(na.rm = TRUE) +
    geom_text(aes(label = percent(val, accuracy = 1)), 
              color = main_gray, size = 5,
              vjust = -1, na.rm = TRUE) +
    scale_y_continuous(label = percent) +
    scale_fill_manual(values=alpha(c(color))) +
    expand_limits(y = 1) +
    facet_wrap(. ~ state) +
    labs(x = "", y = "", subtitle = sex_type) +
    theme_bw() +
    theme(text = element_text(size = 18),
          legend.position = "none",
          axis.title = element_text(color = main_gray),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(fill = "white", color = "white"),
          strip.text = element_text(color = main_gray))
}

grph_cas("Total", comp_teal)

```

NOTE: For 90-90-90 achievement, "[s]ample size is small and should be interpreted with caution."

## We can also see how this looks for females...

```{r cascade_female, fig.width= 16}
grph_cas("Female", main_blue)
```

NOTE: For 90-90-90 achievement, "[s]ample size is small and should be interpreted with caution."

## ...and actual we see higher achievement for males in many states.

```{r cascade_male, fig.width= 16}
grph_cas("Male", comp_purp)
  
  rm(datapack_sex, viz_cas, grph_cas)
```

NOTE: For 90-90-90 achievement, "[s]ample size is small and should be interpreted with caution."


## Comparing program results to NAIIS, Rivers, Egunu and Anambra are heavily under reporting in MER compared with FCT, Adamawa, and Niger which are heavily under reporting.

```{r tx_curr, message=FALSE, warning=FALSE}
  
  tx_curr_est <- cascade_est %>%
  filter(sex == "total") %>% 
  mutate(PLHIV = pop * `Prev 15+`,
         Diagnosed = PLHIV * Diagnosed,
         `On Treatment` = Diagnosed * `On Treatment`) %>% 
  select(state, Diagnosed)

viz_tx_curr_comp <- tx_curr %>% 
  filter(sex == "total") %>% 
  select(-sex) %>% 
  left_join(tx_curr_est, ., by = c("state")) %>%
  arrange(Diagnosed) %>% 
  mutate(state = as_factor(state), 
         reporting = ifelse(Diagnosed > TX_CURR, "MER total under reported", "MER total over reported")) %>%
  gather(type, val, -state, -reporting) %>% 
  arrange(state, type) %>% 
  mutate(lab = case_when(state == "Rivers" & type == "TX_CURR"  ~ "MER",
                         state == "Rivers"& type == "Diagnosed" ~ "NAIIS")) %>% 
  filter(!is.na(reporting))


#viz
viz_tx_curr_comp %>%
  ggplot(aes(state, val)) +
  geom_hline(yintercept = 1, color = comp_gray) +
  geom_path(color = comp_gray, na.rm = TRUE) +
  geom_point(color = ifelse(viz_tx_curr_comp$type == "TX_CURR", comp_gray, main_blue),
             size = 5, na.rm = TRUE) +
  geom_text(aes(label = lab),
            hjust = ifelse(viz_tx_curr_comp$type == "TX_CURR", 1.5, -.3), #vjust = -1,
            size = 5,
            color = ifelse(viz_tx_curr_comp$type == "TX_CURR", comp_gray, main_blue),
            na.rm = TRUE) +
  scale_y_continuous(label = comma) +
  coord_flip() +
  expand_limits(y = 0) +
  labs(x = "", y  = "") + #caption = "Note: Currently missing adult prevalence data by sex required for calculation.") +
  facet_grid(reporting ~ ., scales = "free_y") +
  theme_bw() +
  theme(text = element_text(size = 18),
        legend.position = "none",
        axis.title = element_text(color = main_gray),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        strip.background = element_rect(fill = "white", color = "white"),
        strip.text = element_text(color = main_gray),
        plot.caption = element_text(color = main_gray))
  
  rm(tx_curr_est, viz_tx_curr_comp, cascade_est, reporting)
```

## recap the Discussion Points

- What can we expect post NAIIS based on other country experiences? 
- How to interpret the current Nigeria NAIIS results? What do they mean and what don't they mean? 
- What should be USAID’s focus? Geographic, priority populations, population characteristics, service delivery models, targets, budgets?

## Data Sources

**NOTE: NAIIS data is embargoed and unvalidated at this point**

**NOTE: For 90-90-90 achievement, "[s]ample size is small and should be interpreted with caution."**

- NAIIS data scraped from: CDC. (2018 Oct 18). 2018 Nigeria AIDS Indicator and Impact Suvery (NAIIS):  Preliminary Data, Web 1 States [presentation]
- Spectrum & NAIIS Prevalence comparions extracted from: Delahatu(2018 Oct 19) Comparison of State HIV Prevalence Oct 19 2018_Sebastian.xlsx. CDC.
- MER data extracted from: PEPFAR. (2018, Nov 15). MER Structured Dataset FY18Q4 initial. www.pepfar-panorama.org
- Population and PLHIV estimated from: Nigeria PEPFAR Team. (2018, Mar 16). NigeriaCOP18DataPackv2018.02.13 - Version 10 Mar 16 2018_Final


[Link to Google Drive with data sets](https://drive.google.com/file/d/17SbhzETzdU4GuU198sUqA65S8R5wsmV7/view?usp=sharing)