---
title: "Prevalence in Nigeria"
author: "Aaron Chafetz"
date: "December 6, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.width=12, fig.height=6)

#dependencies
library(tidyverse)
library(readxl)
library(scales)

#colors

main_blue <- "#6585CF"
main_gray <- "#4D4D4D"
comp_gray <- "#808080"
comp_teal <- "#6BB1C9"
comp_purp <- "#7E6BC9"
goal_gold <- "#CEB966"

```

```{r import}

#import
folderpath <- "C:/Users/achafetz/Documents/GitHub/DeepDive/Output"
prev <- read_rds(file.path(folderpath, "prev_surveys.rds"))
burden <- read_rds(file.path(folderpath, "nga_buden.rds"))
datapack <- read_rds(file.path(folderpath, "datapack.rds"))
cascade <- read_rds(file.path(folderpath, "cascade.rds"))
tx_curr <- read_rds(file.path(folderpath, "tx_curr.rds"))
naiis <- read_rds(file.path(folderpath, "naiis.rds"))

```

## Where is NAIIS?

- Preliminary results for NAIIS survey was made available in 6 states.
   - Bauchi, Cross River, Enugu, Kano, Lagos, Nasarawa
- Full results will be released in March



**NOTE: NAIIS data is embargoed and unvalidated at this point**

## Discussion Points to consider

- What can we expect post NAIIS based on other country experiences? 
- How to interpret the current Nigeria NAIIS results? What do they mean and what don't they mean? 
- What should be USAID’s focus? Geographic, priority populations, population characteristics, service delivery models, targets, budgets?

## The initial results show a decline in 23 of the 27 states compared with recent Spectrum prevalence.

```{r prev}

  viz_prev <- prev %>% 
    select(-lab_state) %>% 
    spread(survey, prevalence) %>% 
    mutate(change = Spectrum2018 - NAIIS2018) %>% 
    arrange(desc(change)) %>% 
    mutate(state = as_factor(state)) 
  
  grwth_states <- viz_prev %>%
    mutate(state = as.character(state)) %>% 
    filter(change < 0) %>% 
    pull(state)
  
  viz_prev <- viz_prev %>%
    select(-change) %>% 
    gather(survey, prevalence, -state) %>% 
    mutate(col = case_when(state %in% grwth_states ~ comp_gray,
                           TRUE                    ~ main_blue),
           survey = str_replace(survey, "pectrum20|AIIS20", "'"),
           survey = factor(survey, levels = c("S'16", "S'18", "N'18")))

  viz_prev %>% 
      ggplot(aes(survey, prevalence, group = state)) +
      geom_path(size = 2, color = viz_prev$col) +
      geom_point(size = 6, pch=21, color = "white", fill = viz_prev$col, stroke = 2) +
      scale_y_continuous(#labels=function(x) paste0(x *100,"%"), 
                         breaks = seq(0, .10, .05)) +
      labs(x = "", caption = "S = Spectrum, N = NAIIS") +
      facet_wrap(. ~ state) +
      expand_limits(y = .1) +
      theme_bw() +
      theme(legend.position = "none",
            text = element_text(size = 18),
            axis.ticks = element_blank(),
            axis.title = element_text(color = main_gray),
            panel.border = element_blank(),
            axis.line =  element_blank(),
            strip.background = element_blank(),
            plot.caption = element_text(color = main_gray)) 
  
  rm(viz_prev, grwth_states)
```

## Here is where Nigeria was with PLHIV (Data Pack COP 18)

```{r plhiv}

  viz_plhiv <- burden %>% 
    select(state, plhiv_old, plhiv_est) %>%
    arrange(plhiv_old) %>% 
    mutate(state = as_factor(state)) %>% 
    gather(pd, val, -state) %>% 
    mutate(lab = case_when(state == "Anambra" & pd == "plhiv_old" ~ "Old",
                           state == "Anambra" & pd == "plhiv_est" ~ "NAIIS",
                          TRUE                                 ~ as.character(NA)))

  viz_plhiv_i <- viz_plhiv %>%
    filter(pd == "plhiv_old")
  
  viz_plhiv_i %>% 
    ggplot(aes(state, val)) +
    geom_point(color = ifelse(viz_plhiv_i$pd == "plhiv_old", comp_gray, main_blue),
               size = 6) +
    geom_text(aes(label = lab), 
              hjust = .5, vjust = -1, 
              check_overlap = TRUE,
              size = 5,
              color = ifelse(viz_plhiv_i$pd == "plhiv_old", comp_gray, main_blue),
              na.rm = TRUE) +
    scale_y_continuous(limits = c(0, 300000), label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(y = "PLHIV estimate", x = "") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 18),
          axis.ticks = element_blank(),
          axis.title = element_text(color = main_gray),
          panel.border = element_blank(),
          axis.line = element_line(color = main_gray))
  
  rm(viz_plhiv_i)

```

## ... and now NAIIS projects lower PLHIV in 19 of the 27 initial states.

```{r plhiv_updated}

  viz_plhiv %>% 
    ggplot(aes(state, val)) +
    geom_path(color = main_blue,
              size = 1) +
    geom_point(color = ifelse(viz_plhiv$pd == "plhiv_old", comp_gray, main_blue),
               size = 6) +
    geom_text(aes(label = lab), 
              hjust = .5, vjust = -1, 
              check_overlap = TRUE,
              size = 5,
              color = ifelse(viz_plhiv$pd == "plhiv_old", comp_gray, main_blue),
              na.rm = TRUE) +
    scale_y_continuous(limits = c(0, 300000), label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(y = "PLHIV estimate", x = "") +
    theme_bw() +
    theme(legend.position = "none",
          text = element_text(size = 18),
          axis.ticks = element_blank(),
          axis.title = element_text(color = main_gray),
          panel.border = element_blank(),
          axis.line = element_line(color = main_gray))
  
  rm(viz_plhiv)
```

## Looking at those states in comparison to other states' PLHIV

```{r plhiv_across}
viz_datapack <- datapack %>% 
  arrange(plhiv_old) %>% 
  mutate(state = as_factor(state),
         new_col = ifelse(state %in% unique(prev$state), main_blue, comp_gray))

new_plhiv <- burden %>% 
  mutate(plhiv = prevalence_naiis * pop_num) %>% 
  select(state, plhiv)

suppressWarnings(
viz_datapack <- viz_datapack %>% 
  left_join(new_plhiv, by = "state") %>% 
  arrange(plhiv_old) %>% 
  mutate(state = as_factor(state),
         plhiv_merge = ifelse(!is.na(plhiv), plhiv, plhiv_old),
         placeholder = ifelse(!is.na(plhiv), plhiv_old, as.double(NA))) 
)

viz_datapack %>% 
  ggplot(aes(state, plhiv_old, fill = new_col)) +
  geom_col(fill = viz_datapack$new_col) +
  geom_errorbar(aes(ymax = placeholder, ymin = placeholder), 
                width = .75, size = 1, 
                color = main_blue, 
                na.rm = TRUE) +
  scale_y_continuous(limits = c(0, 300000), labels = comma) +
  labs(x = "", y = "PLHIV", caption = "Gray bars denote states not included in NAIIS") +
  coord_flip() +
  theme_bw() +
  theme(text = element_text(size = 18),
        axis.title = element_text(color = main_gray),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        plot.caption = element_text(color = main_gray))

```

## ...and then the updated PLHIV estimate with NAIIS

```{r plhiv_across_update}

  grwth_states <- viz_prev %>%
    mutate(state = as.character(state)) %>% 
    filter(change < 0) %>% 
    pull(state)

viz_datapack %>% 
  ggplot(aes(state, plhiv_merge, fill = new_col)) +
  geom_col(fill = viz_datapack$new_col) +
  geom_errorbar(aes(ymax = placeholder, ymin = placeholder), 
                width = .75, size = 1, 
                color = ifelse(viz_datapack$placeholder < viz_datapack$plhiv_merge,"white",main_blue), 
                na.rm = TRUE) +
  scale_y_continuous(limits = c(0, 300000), labels = comma) +
  labs(x = "", y = "PLHIV", caption = "Gray bars denote states not included in NAIIS") +
  coord_flip() +
  theme_bw() +
  theme(text = element_text(size = 18),
        axis.title = element_text(color = main_gray),
        axis.ticks = element_blank(),
        panel.border = element_blank())

rm(viz_datapack, new_plhiv)

```

## Given this update, Nigeria still has a large unmet need in these districts


```{r burden}
#unmet need
  viz_burden <- burden %>% 
    select(state, TX_CURR, unmet_burden) %>% 
    arrange(unmet_burden) %>% 
    filter(!is.na(TX_CURR)) %>% 
    mutate(state = as_factor(state)) %>% 
    gather(type, val, -state) %>% 
    arrange(state, type) %>% 
    mutate(col_type = ifelse(type == "TX_CURR", comp_gray, main_blue),
           type = ifelse(type == "unmet_burden", "Unmet Need", type),
           type = as_factor(type))

  viz_burden %>% 
    ggplot(aes(state, val)) +
    geom_col(fill = viz_burden$col_type) +
    scale_y_continuous(label = comma) +
    coord_flip() +
    labs(x = "", y = "PLHIV") +
    theme_bw() +
    theme(text = element_text(size = 18),
          axis.title = element_text(color = main_gray),
          axis.ticks = element_blank(),
          panel.border = element_blank())
  
```

## ... and shows Nigeria still has a ways to go before reaching the second 90.

```{r burden_share}
viz_burden %>% 
    ggplot(aes(state, val, fill = type)) +
    geom_bar(position = "fill",stat = "identity", fill = viz_burden$col_type) +
    geom_hline(linetype = "dashed", yintercept = .81, color = goal_gold, size = 1) +
    scale_y_continuous(label = percent) +
    coord_flip() +
    labs(x = "", y = "Share of PLHIV") +
    theme_bw() +
    theme(text = element_text(size = 18),
          axis.title = element_text(color = main_gray),
          axis.ticks = element_blank(),
          panel.border = element_blank())

  
  rm(viz_burden)
```

## It doesn't appear that our targets heavily align with this unmet need.

```{r burden_mer}
  viz_ind <- burden %>% 
    arrange(unmet_burden) %>% 
    mutate(state = as_factor(state)) %>% 
    select(state, `Unmet Burden` = unmet_burden, TX_NEW, HTS_TST) %>% 
    gather(type, val, -state) %>% 
    mutate(type = factor(type, c("Unmet Burden", "HTS_TST", "TX_NEW")))

  viz_ind %>% 
    ggplot(aes(state, val)) +
    geom_col(fill = ifelse(viz_ind$type == "Unmet Burden", main_blue, comp_teal)) +
    coord_flip() +
    facet_grid(. ~ type, scales = "free_x") +
    scale_y_continuous(label = comma) +
    labs(x = "", y = "") +
    theme_bw() +
    theme(text = element_text(size = 18),
          axis.title = element_text(color = main_gray),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(fill = "white", color = "white"),
          strip.text = element_text(color = main_gray))
  
  rm(viz_ind)
```

## A quick look at the estimated cascade shows low 1st 90, but sucessful 2nd and 3rd 90.

```{r cascade, fig.width= 16}

  datapack_sex <- datapack %>% 
    select(-plhiv_old) %>% 
    gather(sex, pop, -state) %>% 
    mutate(sex = case_when(sex == "pop_num_f" ~ "female",
                           sex == "pop_num_m" ~ "male",
                           TRUE               ~ "total"))
  
  cascade_est <- left_join(naiis, datapack_sex, by = c("state", "sex"))
  
  # viz_cas <- cascade_est %>%
  #   mutate(PLHIV = pop * `Prev 15+`,
  #          Diagnosed = PLHIV * Diagnosed,
  #          `On Treatment` = Diagnosed * `On Treatment`,
  #          `Virally Supressed` = `On Treatment` * `Virally Supressed`) %>% 
  #   select(-`Prev 15+`, -pop, -PLHIV) %>% 
  #   gather(type, val, -state, -sex)
  # 
  
  viz_cas <- cascade_est %>%
    mutate(`On Treatment` = Diagnosed * `On Treatment`,
           `Virally Supressed` = `On Treatment` * `Virally Supressed`) %>% 
    select(-`Prev 15+`, -pop) %>% 
    gather(type, val, -state, -sex) %>% 
    mutate(type = ifelse(type == "On Treatment", "On Tx", type),
           type = case_when(type == "Diagnosed" ~ "1st 90",
                            type == "On Tx"     ~ "2nd 90",
                            type == "Virally Supressed" ~ "3rd 90"),
           sex = str_to_title(sex),
           sex = factor(sex, levels = c("Total", "Female", "Male")))
  
  #order
  state_order <- viz_cas %>% 
    filter(sex == "Total", type == "1st 90") %>% 
    arrange(desc(val)) %>% 
    pull(state)
  
  viz_cas <- viz_cas %>% 
    mutate(state = factor(state, levels = state_order))
  
  #viz
  viz_cas %>% 
    ggplot(aes(type, val, fill = sex)) +
    geom_col() +
    geom_text(aes(label = percent(val, accuracy = 1)), 
              color = main_gray, size = 5,
              vjust = -1) +
    scale_y_continuous(label = percent) +
    scale_fill_manual(values=alpha(c(comp_teal, main_blue, comp_purp))) +
    expand_limits(y = .75) +
    facet_grid(sex ~ state, switch = "y") +
    labs(x = "", y = "") +
    theme_bw() +
    theme(text = element_text(size = 18),
          legend.position = "none",
          axis.title = element_text(color = main_gray),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(fill = "white", color = "white"),
          strip.text = element_text(color = main_gray))
  
  rm(datapack_sex, viz_cas)
```

NOTE: For 90-90-90 achievement, "[s]ample size is small and should be interpreted with caution."

## Comparing program results to NAIIS, Enugu and Cross River are heavily under reporting in MER.

```{r tx_curr}
  
  tx_curr_est <- cascade_est %>%
    mutate(PLHIV = pop * `Prev 15+`,
           Diagnosed = PLHIV * Diagnosed,
           `On Treatment` = Diagnosed * `On Treatment`) %>% 
    select(state:Diagnosed)
  
  viz_tx_curr_comp <- left_join(tx_curr_est, tx_curr, by = c("state", "sex")) %>% 
    gather(type, val, -state, -sex) %>% 
    arrange(state, sex, type) %>% 
    mutate(sex = str_to_title(sex),
           sex = factor(sex, c("Total", "Female", "Male")),
           lab = case_when(state == "Enugu" & type == "TX_CURR" & sex == "Total" ~ "MER",
                           state == "Enugu"& type == "Diagnosed" & sex == "Total" ~ "NAIIS"))
  
  #order
  state_order <- viz_tx_curr_comp %>% 
    filter(sex == "Total", type == "Diagnosed") %>% 
    arrange(val) %>% 
    pull(state)
  
  viz_tx_curr_comp <- viz_tx_curr_comp %>% 
    mutate(state = factor(state, levels = state_order))
  
  #viz
  viz_tx_curr_comp %>%
    ggplot(aes(state, val)) +
    geom_path(color = main_blue) +
    geom_point(color = ifelse(viz_tx_curr_comp$type == "TX_CURR", comp_gray, main_blue),
               size = 5) +
    geom_hline(yintercept = 1, color = comp_gray) +
    geom_text(aes(label = lab),
              hjust = .5, vjust = -1,
              size = 5,
              color = ifelse(viz_tx_curr_comp$type == "TX_CURR", comp_gray, main_blue),
              na.rm = TRUE) +
    scale_y_continuous(label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(x = "", y  = "") +
    facet_wrap(. ~ sex) +
    theme_bw() +
    theme(text = element_text(size = 18),
          legend.position = "none",
          axis.title = element_text(color = main_gray),
          axis.ticks = element_blank(),
          panel.border = element_blank(),
          strip.background = element_rect(fill = "white", color = "white"),
          strip.text = element_text(color = main_gray)) 
  
  rm(tx_curr_est, viz_tx_curr_comp, cascade_est)
```

## recap the Discussion Points

- What can we expect post NAIIS based on other country experiences? 
- How to interpret the current Nigeria NAIIS results? What do they mean and what don't they mean? 
- What should be USAID’s focus? Geographic, priority populations, population characteristics, service delivery models, targets, budgets?

## Data Sources

**NOTE: NAIIS data is embargoed and unvalidated at this point**

**NOTE: For 90-90-90 achievement, "[s]ample size is small and should be interpreted with caution."**

- NAIIS data scraped from: CDC. (2018 Oct 18). 2018 Nigeria AIDS Indicator and Impact Suvery (NAIIS):  Preliminary Data, Web 1 States [presentation]
- Spectrum & NAIIS Prevalence comparions extracted from: Delahatu(2018 Oct 19) Comparison of State HIV Prevalence Oct 19 2018_Sebastian.xlsx. CDC.
- MER data extracted from: PEPFAR. (2018, Nov 15). MER Structured Dataset FY18Q4 initial. www.pepfar-panorama.org
- Population and PLHIV estimated from: Nigeria PEPFAR Team. (2018, Mar 16). NigeriaCOP18DataPackv2018.02.13 - Version 10 Mar 16 2018_Final


[Link to Google Drive with data sets](https://bit.ly/2ETfVOm)