---
title: "Prevalence in Nigeria"
author: "Aaron Chafetz"
date: "December 4, 2018"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#dependencies
library(tidyverse)
library(readxl)
library(scales)
```

```{r import, echo=FALSE}

#import
folderpath <- "C:/Users/achafetz/Documents/GitHub/DeepDive/Output"
prev <- read_rds(file.path(folderpath, "prev_surveys.rds"))
burden <- read_rds(file.path(folderpath, "nga_buden.rds"))
datapack <- read_rds(file.path(folderpath, "datapack.rds"))
cascade <- read_rds(file.path(folderpath, "cascade.rds"))
tx_curr <- read_rds(file.path(folderpath, "tx_curr.rds"))
naiis <- read_rds(file.path(folderpath, "naiis.rds"))

```

## Where is NAIIS?

- Preliminary results for NAIIS survey was made available in 6 states.
   - Bauchi, Cross River, Enugu, Kano, Lagos, Nasarawa
- Full results will be released in March



**NOTE: NAIIS data is embargoed and unvalidated at this point**

## Discussion Points to consider

- What can we expect post NAIIS based on other country experiences? 
- How to interpret the current Nigeria NAIIS results? What do they mean and what don't they mean? 
- What should be USAID’s focus? Geographic, priority populations, population characteristics, service delivery models, targets, budgets?

## The initial results show a decline in 5 of the six states compared with recent Spectrum prevalence.

```{r prev, warning = FALSE, echo=FALSE}
  prev %>% 
    ggplot(aes(survey, prevalence, group = state, color = state)) +
    geom_path() +
    geom_point() +
    geom_text(aes(label = lab_state), hjust = 1.1, vjust = .3, na.rm = TRUE) +
    scale_y_continuous(label = percent) +
    labs(x = "") +
    theme(legend.position = "none")
```

## Here is where Nigeria was with PLHIV (Data Pack COP 18)

```{r plhiv, warning = FALSE, echo=FALSE}

  viz_plhiv <- burden %>% 
    select(state, plhiv_old, plhiv_est) %>%
    arrange(plhiv_old) %>% 
    mutate(state = as_factor(state)) %>% 
    gather(pd, val, -state) %>% 
    mutate(lab = case_when(state == "Kano" & pd == "plhiv_old" ~ "Old",
                             state == "Kano" & pd == "plhiv_est" ~ "NAIIS",
                             TRUE                                 ~ as.character(NA)))

  viz_plhiv_i <- viz_plhiv %>%
    filter(pd == "plhiv_old")
  
  viz_plhiv_i %>% 
    ggplot(aes(state, val)) +
    geom_point(color = ifelse(viz_plhiv_i$pd == "plhiv_old", "#808080", "#000000"),
               size = 5) +
    geom_text(aes(label = lab), 
              hjust = .5, vjust = -1, 
              check_overlap = TRUE,
              color = ifelse(viz_plhiv_i$pd == "plhiv_old", "#808080", "#000000"),
              na.rm = TRUE) +
    scale_y_continuous(label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(y = "PLHIV estimate",
         x = "")
  
  rm(viz_plhiv_i)

```

## ... and now NAIIS projects lower PLHIV in five of the six initial states.

```{r plhiv_updated, warning = FALSE, echo=FALSE}

  viz_plhiv %>% 
    ggplot(aes(state, val)) +
    geom_path(color = "#808080") +
    geom_point(color = ifelse(viz_plhiv$pd == "plhiv_old", "#808080", "#000000"),
               size = 5) +
    geom_text(aes(label = lab), 
              hjust = .5, vjust = -1, 
              check_overlap = TRUE,
              color = ifelse(viz_plhiv$pd == "plhiv_old", "#808080", "#000000"),
              na.rm = TRUE) +
    scale_y_continuous(label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(y = "PLHIV estimate",
         x = "")
  
  rm(viz_plhiv_i)
```

## Looking at those states in comparison to other states' PLHIV

```{r plhiv_across, warning = FALSE, echo=FALSE}
viz_datapack <- datapack %>% 
  arrange(plhiv_old) %>% 
  mutate(state = as_factor(state),
         new_col = ifelse(state %in% unique(prev$state), "orange", "gray"))

viz_datapack %>% 
  ggplot(aes(state, plhiv_old, fill = new_col)) +
  geom_col(fill = viz_datapack$new_col) +
  scale_y_continuous(labels = comma) +
  labs(x = "", y = "PLHIV") +
  coord_flip()

```

## ...and then the updated PLHIV estimate with NAIIS

```{r plhiv_across_update, warning = FALSE, echo=FALSE}

new_plhiv <- burden %>% 
  mutate(plhiv = prevalence_naiis * pop_num) %>% 
  select(state, plhiv)

viz_datapack2 <- viz_datapack %>% 
  left_join(new_plhiv, by = "state") %>% 
  arrange(plhiv_old) %>% 
  mutate(state = as_factor(state),
         plhiv_merge = ifelse(!is.na(plhiv), plhiv, plhiv_old),
         placeholder = ifelse(!is.na(plhiv), plhiv_old, as.double(NA))) 

viz_datapack2 %>% 
  ggplot(aes(state, plhiv_merge, fill = new_col)) +
  geom_col(fill = viz_datapack2$new_col) +
  geom_errorbar(aes(ymax = placeholder, warning = FALSE, ymin = placeholder), width = .75, size = 1, color = "black", na.rm = TRUE) +
  scale_y_continuous(labels = comma) +
  labs(x = "", y = "PLHIV") +
  coord_flip()

rm(viz_datapack, viz_datapack2, new_plhiv)

```

## Given this update, Nigeria still has a large unmet need in these districts


```{r burden, warning = FALSE, echo=FALSE}
#unmet need
  viz_burden <- burden %>% 
    select(state, TX_CURR, unmet_burden) %>% 
    arrange(unmet_burden) %>% 
    mutate(state = as_factor(state)) %>% 
    gather(type, val, -state) %>% 
    arrange(state, desc(type)) %>% 
    mutate(type = ifelse(type == "unmet_burden", "Unmet Need", type),
           type = as_factor(type))

  viz_burden %>% 
    ggplot(aes(state, val, fill = type)) +
    geom_col() +
    scale_y_continuous(label = comma) +
    coord_flip() +
    labs(x = "", y = "PLHIV")
  
```

## ... and shows Nigeria still has a ways to go before reaching the second 90.

```{r burden_share, warning = FALSE, echo=FALSE}
viz_burden %>% 
    ggplot(aes(state, val, fill = type)) +
    geom_bar(position = "fill",stat = "identity") +
    geom_hline(yintercept = .81, color = "#808080", size = 2) +
    scale_y_continuous(label = percent) +
    coord_flip() +
    labs(x = "", y = "Share of PLHIV")

  
  rm(viz_burden)
```

## It doesn't appear that our targets heavily align with this unmet need.

```{r burden_mer, warning = FALSE, echo=FALSE}
  viz_ind <- burden %>% 
    arrange(unmet_burden) %>% 
    mutate(state = as_factor(state)) %>% 
    select(state, `Unmet Burden` = unmet_burden, TX_NEW, HTS_TST) %>% 
    gather(type, val, -state) %>% 
    mutate(type = factor(type, c("Unmet Burden", "HTS_TST", "TX_NEW")))

  viz_ind %>% 
    ggplot(aes(state, val)) +
    geom_col(fill = ifelse(viz_ind$type == "Unmet Burden", "gray", "#808080")) +
    coord_flip() +
    facet_grid(. ~ type, scales = "free_x") +
    scale_y_continuous(label = comma) +
    labs(x = "", y = "")
  
  rm(viz_ind)
```

## A quick look at the estimated cascade shows low 1st 90, but sucessful 2nd and 3rd 90.

```{r cascade, warning = FALSE, echo=FALSE}

  datapack_sex <- datapack %>% 
    select(-plhiv_old) %>% 
    gather(sex, pop, -state) %>% 
    mutate(sex = case_when(sex == "pop_num_f" ~ "female",
                           sex == "pop_num_m" ~ "male",
                           TRUE               ~ "total"))
  
  cascade_est <- left_join(naiis, datapack_sex, by = c("state", "sex"))
  
  # viz_cas <- cascade_est %>%
  #   mutate(PLHIV = pop * `Prev 15+`,
  #          Diagnosed = PLHIV * Diagnosed,
  #          `On Treatment` = Diagnosed * `On Treatment`,
  #          `Virally Supressed` = `On Treatment` * `Virally Supressed`) %>% 
  #   select(-`Prev 15+`, -pop, -PLHIV) %>% 
  #   gather(type, val, -state, -sex)
  # 
  
  viz_cas <- cascade_est %>%
    mutate(`On Treatment` = Diagnosed * `On Treatment`,
           `Virally Supressed` = `On Treatment` * `Virally Supressed`) %>% 
    select(-`Prev 15+`, -pop) %>% 
    gather(type, val, -state, -sex) %>% 
    mutate(type = ifelse(type == "On Treatment", "On Tx", type))
  
  viz_cas %>% 
    ggplot(aes(type, val)) +
    geom_col() +
    geom_text(aes(label = percent(val, accuracy = 1)), vjust = -1) +
    scale_y_continuous(label = percent) +
    expand_limits(y = .75) +
    facet_grid(sex ~ state) +
    labs(x = "", y = "") +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank())
  
  rm(datapack_sex, viz_cas)
```

## Comparing program results to NAIIS, Enugu and Cross River are heavily under reporting in MER.

```{r tx_curr, warning = FALSE, echo=FALSE}
  
  tx_curr_est <- cascade_est %>%
    mutate(PLHIV = pop * `Prev 15+`,
           Diagnosed = PLHIV * Diagnosed,
           `On Treatment` = Diagnosed * `On Treatment`) %>% 
    select(state:Diagnosed)
  
  viz_tx_curr_comp <- left_join(tx_curr_est, tx_curr, by = c("state", "sex")) %>% 
    gather(type, val, -state, -sex) %>% 
    arrange(state, sex, type)
  
  viz_tx_curr_comp %>%
    ggplot(aes(state, val)) +
    geom_path(color = "#808080") +
    geom_point(color = ifelse(viz_tx_curr_comp$type == "TX_CURR", "#808080", "#000000"),
               size = 5) +
    # geom_text(aes(label = lab), 
    #           hjust = .5, vjust = -1, 
    #           check_overlap = TRUE,
    #           color = ifelse(viz_tx_curr_comp$type == "TX_CURR", "#808080", "#000000"),
    #           na.rm = TRUE) +
    scale_y_continuous(label = comma) +
    coord_flip() +
    expand_limits(y = 0) +
    labs(x = "", y  = "") +
    facet_wrap(. ~ sex) +
    theme(legend.position = "top") 
  
  rm(tx_curr_est, viz_tx_curr_comp, cascade_est)
```

## recap the Discussion Points

- What can we expect post NAIIS based on other country experiences? 
- How to interpret the current Nigeria NAIIS results? What do they mean and what don't they mean? 
- What should be USAID’s focus? Geographic, priority populations, population characteristics, service delivery models, targets, budgets?

## Data Sources

**NOTE: NAIIS data is embargoed and unvalidated at this point**

- NAIIS data scraped from: CDC. (2018 Oct 18). 2018 Nigeria AIDS Indicator and Impact Suvery (NAIIS):  Preliminary Data, Web 1 States [presentation]
- Spectrum & NAIIS Prevalence comparions extracted from: Delahatu(2018 Oct 19) Comparison of State HIV Prevalence Oct 19 2018_Sebastian.xlsx. CDC.
- MER data extracted from: PEPFAR. (2018, Nov 15). MER Structured Dataset FY18Q4 initial. www.pepfar-panorama.org
- Population and PLHIV estimated from: Nigeria PEPFAR Team. (2018, Mar 16). NigeriaCOP18DataPackv2018.02.13 - Version 10 Mar 16 2018_Final


[Link to Google Drive with data sets](https://bit.ly/2ETfVOm)